---
title: "295_project"
author: "Ali javed"
date: "11/2/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
library(ISLR)
setwd('/Users/alijaved/Desktop/CS295_ISLR/project_git/295-final-master')
SP500 = read.csv("SP500.csv")  # read csv file 
NASDAQ = read.csv("NASDAQCOM.csv")
HAPPINESS = read.csv("sumhapps.csv")
colnames(HAPPINESS) <- c("DATE", "HAPPINESS")


#convert date to date format
SP500[[1]] <- as.Date(SP500[[1]])
NASDAQ[[1]] <- as.Date(NASDAQ[[1]])
HAPPINESS[[1]] <- as.Date(HAPPINESS[[1]])

#convert index value to numeric
SP500$SP500<- as.numeric(as.character(SP500$SP500))
NASDAQ$NASDAQCOM<- as.numeric(as.character(NASDAQ$NASDAQCOM))
HAPPINESS$HAPPINESS<- as.numeric(as.character(HAPPINESS$HAPPINESS))

#create binary columns for increase and decrease
#take difference with last days value
SP500$SP500_difference <- c(0, diff(SP500$SP500))
NASDAQ$NASDAQCOM_difference <- c(0, diff(NASDAQ$NASDAQCOM))
HAPPINESS$HAPPINESS_difference <-c(0,diff(HAPPINESS$HAPPINESS))

#make binary direction
SP500$SP500_direction <- ifelse(SP500$SP500_difference>0, 1,0)
NASDAQ$NASDAQCOM_direction <- ifelse(NASDAQ$NASDAQCOM_difference>0, 1,0)
HAPPINESS$HAPPINESS_direction <- ifelse(HAPPINESS$HAPPINESS_difference>0, 1,0)

#merge SP500, NASDAQ and HAPPINESS into single table
data =merge(SP500,NASDAQ, by ='DATE')
data = merge(data, HAPPINESS, by = 'DATE')
data$day <- weekdays(as.Date(data$DATE))


#remove seperate tables
rm(SP500)
rm(HAPPINESS)
rm(NASDAQ)
```

Initial analysis. Pairs and coorelation.



```{r}
pairs(data[,c("SP500","NASDAQCOM")])
sub = subset(data,select=-c(day,DATE))
cor(data[,c("SP500","NASDAQCOM")])
```





Lets do a logistic regression. Too much co-relation between both!

```{r}

lm.fit = lm(SP500~NASDAQCOM, data= data)
summary(lm.fit)
```
```{r}
glmSP500.fit = glm(SP500_direction~NASDAQCOM_difference, data= data, family = binomial)
summary(glmSP500.fit)


```



logistic regression plot predicting SP500 direction (up or down) based on the NASDAQ difference from last day

```{r}
newdat <- data.frame(NASDAQCOM_difference=seq(-206, 206,len=200))
newdat$vs = predict(glmSP500.fit, newdata=newdat, type="response")
#plot(vs~hp, data=my_data, col="red4")

newdat$vs<- as.numeric(as.character(newdat$vs))
plot(data$NASDAQCOM_difference,data$SP500_direction)
lines(newdat$vs ~ newdat$NASDAQCOM_difference)

#abline(glm.fit)
```

This plot does show that somtimes when NASDAQ goes up, SP is going down, but if the difference in NASDAQ is really big (in positive direction), SP500 does indeed always go up.

However, it makes more sense to see whether the we can use NASDAQ price a few days earlier to predict SP500 price, or so use SP500s price a few days earlier to predit its price going forward.

Calculate lag variables

```{r}
library(Hmisc)
#lag for SP500
data$SP500_lag1 <- Lag(data$SP500, +1)
data$SP500_lag2 <- Lag(data$SP500, +2)
data$SP500_lag10 <- Lag(data$SP500, +10)

#lag for NASDAQCOM
data$NASDAQCOM_lag1 <- Lag(data$NASDAQCOM, +1)
data$NASDAQCOM_lag2 <- Lag(data$NASDAQCOM, +2)
data$NASDAQCOM_lag10 <- Lag(data$NASDAQCOM, +10)

#lag for HAPPINESS
data$HAPPINESS_lag1 <- Lag(data$HAPPINESS,+1)

#difference for SP500
data$SP500_lag1_difference <- c(0, diff(data$SP500_lag1))
data$SP500_lag2_difference <- c(0, diff(data$SP500_lag2))
data$SP500_lag10_difference <- c(0, diff(data$SP500_lag10))

#difference for NASDAQCOM
data$NASDAQCOM_lag1_difference <- c(0, diff(data$NASDAQCOM_lag1))
data$NASDAQCOM_lag2_difference <- c(0, diff(data$NASDAQCOM_lag2))
data$NASDAQCOM_lag10_difference <- c(0, diff(data$NASDAQCOM_lag10))

#DIFFERENCE FOR HAPPINESS
data$HAPPINESS_lag1_difference <- c(0, diff(data$HAPPINESS_lag1))

#direction for SP500 lag variables
data$SP500_lag1_direction <- ifelse(data$SP500_lag1_difference>0, 1,0)
data$SP500_lag2_direction <- ifelse(data$SP500_lag2_difference>0, 1,0)
data$SP500_lag10_direction <- ifelse(data$SP500_lag10_difference>0, 1,0)

#direction for NASDAQCOM lag variables
data$NASDAQCOM_lag1_direction <- ifelse(data$NASDAQCOM_lag1_difference>0, 1,0)
data$NASDAQCOM_lag2_direction <- ifelse(data$NASDAQCOM_lag2_difference>0, 1,0)
data$NASDAQCOM_lag10_direction <- ifelse(data$NASDAQCOM_lag10_difference>0, 1,0)

```

Now that we have data, lets see what days are the best to predict direction going forward using coorelation. In the graphs the bottom most row is important, if we can fit a regression model with significant accuracy there, it means that we predict whether or not the index will go up or down, based on how much it went up or down X days ago. 

```{r}
pairs(data[,c("NASDAQCOM_lag1_difference","NASDAQCOM_lag2_difference","NASDAQCOM_lag10_difference","SP500_direction")])

pairs(data[,c("HAPPINESS_difference","SP500_lag10_difference","NASDAQCOM_direction")])
```
```{r}
data_na = na.omit(data)
#cor(data_na[,c("HAPPINESS_difference","SP500_lag1","SP500_lag2","SP500_lag10","SP500")])
pairs(data[,c("SP500_difference","SP500_lag10_difference")])

```

```{r}
## 75% of the sample size
data_na = na.omit(data)
smp_size <- floor(0.75 * nrow(data_na))

## set the seed to make your partition reproductible
set.seed(1)
train_ind <- sample(seq_len(nrow(data_na)), size = smp_size)

train <- data_na[train_ind, ]
test <- data_na[-train_ind, ]


```

```{r}
glmSP500.fit=glm(SP500_direction~SP500_lag1_difference, data= train, family = binomial)
summary(glmSP500.fit)

```
```{r}
glm.probs =predict(glmSP500.fit, test, type="response")
glm_link_scores <- predict(glmSP500.fit, test, type="link")
```


```{r}
glm.probs =predict(glmSP500.fit, test, type="response")
glm.pred=rep(0 ,467)
glm.pred[glm.probs>.5]=1
table(glm.pred ,test$SP500_direction)
mean(glm.pred == test$SP500_direction)
```


```{r}
library(pROC)
plot(roc(test$SP500_direction, glm.probs, direction="<"),
     col="yellow", lwd=3, main="The turtle finds its way")

glm_simple_roc <- simple_roc(test_set$SP500_direction==1, glm_link_scores)
with(glm_simple_roc, points(1 - FPR, TPR, col=1 + labels))
```



```{r}
qdaSP500.fit=glm(SP500_direction~NASDAQCOM, data= train, family = binomial)
summary(qdaSP500.fit)
```


```{r}
qdaSP500.probs =predict(qdaSP500.fit, test, type="response")
qdaSP500.pred=rep(0 ,467)
qdaSP500.pred[glm.probs>.6]=1
table(qdaSP500.pred ,test$SP500_direction)
mean(qdaSP500.pred == test$SP500_direction)

```

```{r}
newdat <- data.frame(SP500_lag10_difference=seq(-106, 106,len=100))
newdat$vs = predict(glmNASDAQCOM.fit_lagSP_lag_10, newdata=newdat, type="response")
#plot(vs~hp, data=my_data, col="red4")

newdat$vs<- as.numeric(as.character(newdat$vs))
plot(data$SP500_lag10_difference,data$NASDAQCOM_direction)
lines(newdat$vs ~ newdat$SP500_lag10_difference)

#abline(glm.fit)
```
```{r}

#data = data[12:nrow(data),]
#set.seed(1)
#data = na.omit(data)
#glm.fit = glm(day~HAPPINESS, data = data)
#forest.model = randomForest(HAPPINESS~.,data,mtry = 1, ntrees = 10)


#bag.carseats= randomForest(Sales~.,Carseats, subset = train,mtry = dim(Carseats)[2] -1, importance = TRUE)



#bag.carseats= randomForest(Sales~.,Carseats, subset = train,mtry = dim(Carseats)[2] -1, importance = TRUE)

#tree.predict = predict(bag.carseats, newdata=test.carseats )

#MSE = mean((tree.predict-test.sales)^2)

#print (MSE)
```
```{r}
library(ISLR)
library(MASS)

qda.fit=qda(day~HAPPINESS+HAPPINESS_difference ,data=train)

qda.fit
```



```{r}
#data.test = data_na[-train,]
qda.class =predict(qda.fit ,train)
t= test$day

#t = data.test$day
#table(qda.class$class ,t)
accuracy(qda.class$class ,t)
#qda.class$class
#mean(qda.class$class == t)

#qda.class$class
```

